---
title: "Quickstart"
eval: false
---

This guide aims to help you get started with Hal9 in your local development environment as quickly as possible.

You can also head over to the hosted demo at [https://hal9.com/new](https://hal9.com/new) without having to download
or install things locally.

### Installation

Currently, to install the package you need a working Rust toolchain, we recommend grabbing one using [rustup.rs](https://rustup.rs) if you don't have one yet.

In addition, if you're installing the R package on *Windows*, you'll need to have [Rtools](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html)
installed.


::: {.panel-tabset group="language"}

## R

```{r}
remotes::install_github("hal9ai/hal9", subdir = "r")
```

## CLI 

```{bash}
cargo install --git https://github.com/hal9ai/hal9.git
```
:::

### Starter apps

The recommended way to bootstrap a new app is to use the `hal9 new` utility:

::: {.panel-tabset group="language"}

## R

```{r}
library(hal9)
h9_new("my_app")
```

## CLI

```{bash}
hal9 new my_app
```

:::

This creates a new directory `my_app/` with the minimal ingredients necessary to define a Hal9 app.

Now that we have an app, we can launch the server and see it in action:

::: {.panel-tabset group="language"}

## R

```{r}
h9_start("my_app")
```

## CLI

```{bash}
hal9 start my_app
```

:::

A browser tab should now pop up, allowing you to make changes to the app.