---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  echo = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hal9

<!-- badges: start -->
<!-- badges: end -->

The goal of hal9 is to provide R users a high-level hal9.js API.

## Installation

You can install the development version of hal9 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hal9ai/hal9ai")
```

## Basic Usage

Hal9 is a javascript library that enables anyone to compose visualizations and predictive models optimized for websites and web APIs.

You can explore your data interactively using the `hal9` function:

```{r example}
library(hal9)
## basic example code

hal9(my_data)
```

You may also build a specific pipeline using our high-level functions:

```{r}

# option 1

# step-by-step construction
hal9_pipeline() |> 
  hal9_add_data(iris) |> 
  hal9_add_filter() |> 
  hal9_show()

#option 2

# pre-built common pipelines
hal9_filter_data(iris)
```

## Exporting a pipeline

You may export an existing pipeline built using high level functions to a html file using the `hal9_render` function:

```{r}

hal9_pipeline() |> 
  hal9_add_data(iris) |> 
  hal9_add_filter() |> 
  hal9_render("pipeline.html")

```

## Publishing a pipeline

You can also publish your work on RPubs and make it easily shareable through a link:

```{r}

hal9_pipeline() |> 
  hal9_add_data(iris) |> 
  hal9_add_filter() |> 
  hal9_publish()

```

# TO DO

- [ ] Test infrastructure
- [ ] `hal9_render()` first draft
- [ ] `hal9_render()` tests
- [ ] `hal9_publish()` first draft
